A parametricity-based formalization of semi-simplicial and semi-cubical sets

### 2023-01-23 paper/paper.tex
paper: fix compile, minor nits

Notice that each $\restrl$ in the type of $\cohl$ is hiding a $\cohf$ rewrite: this makes a sum total of three $\cohf$ rewrites on the left-hand side, and two $\cohf$ rewrites on the right-hand side. In the proof term of $\cohl$, $\cohc$ has one $\cohf$ rewrite on its left-hand side and zero on the right-hand side. This combined with the two terms of the form $\ap \cohf$ matches our expectation of three $\cohf$ on the left-hand side and two $\cohf$ on the right-hand side. $\coht$ can roughly be seen as a commutation of these $\cohf$ terms.

@@

For a fixed $n$, $\framep$, $\restrf$, and $\cohf$ have to be mutually defined by induction on $p$. Each of them rely on $\painting$, $\restrf$, and $\cohc$ at lower $n$, so they are well-founded. Finally, each of $\painting$, $\restrf$, and $\cohf$ are by induction from $p$ to $n$.

Remark: If we weren't working in \U, but in \HGpd, we would need to prove one more higher-dimensional coherence, and if we were working in \Type, we would need to prove arbitrarily many higher-dimensional coherences. For the unfamiliar reader, $\HGpd$ is the subset of types $A$ such that for all $x$ and $y$ in $A$, $x =} y$ is in \U. See \cite{herbelin15,altenkirch16,kraus21} for a discussion on the need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory.

@@

Since the construction shown in the previous sections are by induction on $n$, and dependencies are on lower $n$ and $p < n$, one would imagine formalizing this using well-founded induction in the calculus of inductive constructions. We initially tried this approach: we had terms dependent on the proofs of the case distinction of $n' \leq n$ implies $n' < n$ or $n' = n$, and these proofs didn't have definitional computational rule; it started to be necessary to reason propositionally on the computational property of the case distinction, and it eventually turned out to be unmanageable. Hence, we chose a different route: in practice, since $\restrf[n]$ depends on $\framep[n]$ and $\framep[n-1]$, while $\cohf[n]$ depends on $\framep[n]$, $\framep[n-1]$, and $\framep[n-2]$, we only need to keep track of three ``levels'', and we built separate data structures for the levels, with dependencies. More concretely, we build the ten definitions shown in the tables by induction, and this is part of the definition of a larger record. The other fields of the record are $\framep$, $\layer$, $\painting$ at levels $n - 1$ and $n - 2$, $\restrf$, $\restrl$, and $\restrc$ at level $n - 1$, and equations to recall the definitions of these objects at lower levels.

>>>>

Notice that each $\restrl$ in the type of $\cohl$ is hiding a $\cohf$ rewrite: this makes a sum total of three $\cohf$ rewrites on the left-hand side, and two $\cohf$ rewrites on the right-hand side. In the proof term of $\cohl$, $\cohc$ has one $\cohf$ rewrite on its left-hand side and zero on the right-hand side. This combined with the two terms of the form $\ap \cohf$ matches our expectation of three $\cohf$ on the left-hand side and two $\cohf$ on the right-hand side. Then, $\coht$ can roughly be seen as a commutation of these $\cohf$ terms.

@@

For a fixed $n$, $\framep$, $\restrf$, and $\cohf$ have to be mutually defined by induction on $p$. Each of them rely on $\painting$, $\restrc$, and $\cohc$ at lower $n$, so they are well-founded. Finally, each of $\painting$, $\restrc$, and $\cohc$ are by induction from $p$ to $n$.

Remark: If we weren't working in \U, but in \HGpd, we would need to prove one more higher-dimensional coherence, and if we were working in \Type, we would need to prove arbitrarily many higher-dimensional coherences. For the unfamiliar reader, $\HGpd$ is the subset of types $A$ such that for all $x$ and $y$ in $A$, $x = y$ is in \U. See \cite{herbelin15,altenkirch16,kraus21} for a discussion on the need for recursive higher-dimensional coherence conditions in formulating higher-dimensional structures in type theory.

@@

Since the construction shown in the previous sections is by induction on $n$, and dependencies are on lower $n$ and $p < n$, one would imagine formalizing this using well-founded induction in dependent type theory. We initially tried this approach: we had terms dependent on the proofs of the case distinction of $n' \leq n$ implies $n' < n$ or $n' = n$, and these proofs did not have definitional computational rule; it started to be necessary to reason propositionally on the computational property of the case distinction, and it eventually turned out to be unmanageable. Hence, we chose a different route: in practice, since $\restrf[n]$ depends on $\framep[n]$ and $\framep[n-1]$, while $\cohf[n]$ depends on $\framep[n]$, $\framep[n-1]$, and $\framep[n-2]$, we only need to keep track of three ``levels'', and we built separate data structures for the levels, with dependencies. More concretely, we build the ten definitions shown in the tables by induction, and this is part of the definition of a larger record. The other fields of the record are $\framep$, $\layer$, $\painting$ at levels $n - 1$ and $n - 2$, $\restrf$, $\restrl$, and $\restrc$ at level $n - 1$, and equations to recall the definitions of these objects at lower levels.
